---
title: "Homework 2"
subtitle: "ECON 470, Spring 2025"
author: "Molly Catlin"
format:
  pdf:
    output-file: "catlin_molly_hwk2_fs"
    output-ext: "pdf"
    header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}
---


```{r}
#| include: false
#| eval: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra)
```


Here is a link to my repository: {https://github.com/mollyjc02/Homework_2.git}


\newpage



```{r}
#| include: false

####load("C:/Users/Molly/OneDrive - Emory University/Documents/GitHub/Homework_2/final_submission/Hwk2_workspace.RData")

library(here)
here()

# Load the data using a relative path
load(here("final_submission/Hwk2_workspace.RData")) 
```




\noindent **1. How many hospitals filed more than one report in the same year? Show your answer as a line graph of the number of hospitals over time.**


```{r}
#| echo: false
#| label: duplicate-hospitals
#| tbl-cap: Hospitals that Filed Multiple Reports
library(ggplot2)

ggplot(multi_report_counts, aes(x = fyear, y = num_hospitals)) +
geom_line(color = "blue", linewidth  = 1) +
geom_point(color = "red", size  = 2) +
labs(
title = "Hospitals Filing Multiple Reports Per Year",
x = "Year",
y = "Number of Hospitals",
caption = "Source: HCRIS Data (1996 & 2010 Versions)"
) +
theme_minimal()
```


\newpage

\noindent **2. After removing/combining multiple reports, how many unique hospital IDs (Medicare provider numbers) exist in the data?**


```{r}
#| echo: false
#| label: unique-hospitals 
#| tbl-cap: Unique Hospitals 

print(fig.unique)
```


\newpage

\noindent **3. What is the distribution of total charges (tot_charges in the data) in each year?**


```{r}
#| echo: false
#| label: charge-distributions
#| tbl-cap: Charge Distribution by Year

ggplot(charge.data, aes(x = factor(year), y = tot_charges)) +
  geom_violin(fill = "lightblue", color = "darkblue", alpha = 0.6) +
  scale_y_log10(labels = scales::comma) + 
  labs(
    title = "Distribution of Total Charges by Year",
    x = "Year",
    y = "Total Charges (log scale)",
    caption = "Source: HCRIS Data (1996 & 2010 Versions)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```


\newpage

\noindent **4. What is the distribution of estimated prices in each year?**

```{r}
#| echo: false
#| label: price-distributions
#| tbl-cap: Estimated Price Distribution by Year


ggplot(price.data, aes(x = as.factor(year), y = price)) +
  geom_violin(trim = TRUE, fill = "skyblue", color = "black") +
  labs(
    title = "Distribution of Estimated Prices by Year",
    x = "Year",
    y = "Estimated Price"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```



\newpage

**For the rest of the assignment, I have included only observations in 2012. So I am now dealing with cross-sectional data in which some hospitals are penalized and some are not.**


\noindent **5. Calculate the average price among penalized versus non-penalized hospitals.**

The average price among penalized hospitals is `r mean.pen`, while the average price among non-penalized hospitals is `r mean.nopen`. 

\newpage

\noindent **6. Split hospitals into quartiles based on bed size and provide a table of the average price among treated/control groups for each quartile.**

```{r}
#| echo: false
#| label: price-by-quartile
#| tbl-cap: Average Price by Treatment Status and Bed Size 

knitr::kable(
  quartile_summary,
  col.names = c("Bed Quartile", "No Penalty", "Penalty"),
  caption = "Average Prices by Bed Quartile and Penalty Status",
  format = "markdown"
)
```


\newpage


\noindent **7. Find the average treatment effect based on quartiles of bed size using each of the following estimators: nearest neighbor matching with inverse variance distance, nearest neighbor matching with Mahalanobis distance, inverse propensity weighting, and simple linear regression.**

```{r}
#| echo: false
#| label: ate-table
#| tbl-cap: ATE Estimates

# Extract ATE estimates
ate_near_match <- near.match$est
ate_mal_match <- mal.match$est
ate_ipw_diff <- ipw.diff
ate_regression <- ate


# Create a data frame summarizing the results
ate_estimates <- data.frame(
  Method = c(
    "Nearest Matching (Inverse Variance)",
    "Nearest Matching (Mahalanobis Distance)",
    "Inverse Propensity Weighting (IPW)",
    "Linear Regression"
  ),
  ATE_Estimate = c(ate_near_match, ate_mal_match, ate_ipw_diff, ate_regression)
)


# Print the table using knitr::kable()
knitr::kable(ate_estimates,
             caption = "ATE Estimates from Different Methods",
             col.names = c("Method", "ATE Estimate"))
```


\noindent **8. f**