---
title: "Homework 1"
subtitle: "ECON 470, Spring 2025"
author: "Molly Catlin"
format:
  pdf:
    output-file: "catlin_molly_hwk2_fs"
    output-ext:  "pdf"
    header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}
---

```{r}
#| include: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra)
```

Here is a link to my repository: {https://github.com/mollyjc02/Homework_2.git}

\newpage 

```{r}
#| include: false
#| eval: true

load("C:/Users/Molly/OneDrive - Emory University/Documents/GitHub/Homework_2/final_submission/results/Hwk2_workspace.RData")
```


\noindent 1. How many hospitals filed more than one report in the same year? Show your answer as a line graph of the number of hospitals over time. 

```{r} 
#| echo: false
#| label: tbl-plans
#| tbl-cap: Hospitals that Filed Multiple Reports
library(ggplot2)

ggplot(multi_report_counts, aes(x = fyear, y = num_hospitals)) +
geom_line(color = "blue", size = 1) +
geom_point(color = "red", size = 2) +
labs(
title = "Hospitals Filing Multiple Reports Per Year",
x = "Year",
y = "Number of Hospitals",
caption = "Source: HCRIS Data (1996 & 2010 Versions)"
) +
theme_minimal()
```


\noindent 2. After removing/combining multiple reports, how many unique hospital IDs (Medicare provider numbers) exist in the data?

```{r} 
#| echo: false
#| eval: true

print(fig.unique)
```


\noindent 3. What is the distribution of total charges (tot_charges in the data) in each year?

```{r}
#| echo: false
#| label: tbl-plans
#| tbl-cap: Multiple reports

ggplot(charge.data, aes(x = factor(year), y = tot_charges)) +
geom_violin(fill = "lightblue", color = "darkblue", alpha = 0.6) +
scale_y_log10() + # Apply log scale to handle skewed distributions
labs(
title = "Distribution of Total Charges by Year",
x = "Year",
y = "Total Charges (log scale)",
caption = "Source: HCRIS Data (1996 & 2010 Versions)"
) +
theme_minimal()
```


\noindent 4. What is the distribution of estimated prices in each year?

```{r}
#| echo: false
#| label: tbl-plans
#| tbl-cap: Multiple reports

ggplot(price.data, aes(x = as.factor(year), y = price)) +
  geom_violin(trim = TRUE, fill = "skyblue", color = "black") +
  labs(
    title = "Distribution of Estimated Prices by Year",
    x = "Year",
    y = "Estimated Price"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```


\newpage 

For the rest of the assignment, I have included only observations in 2012. So I am now dealing with cross-sectional data in which some hospitals are penalized and some are not.


\noindent 5. Calculate the average price among penalized versus non-penalized hospitals.


```{r}
#| echo: false
#| label: tbl-plans
#| tbl-cap: Multiple reports

cat("Average price among penalized hospitals: ", mean.pen, "\n")
cat("Average price among non-penalized hospitals: ", mean.nopen, "\n")
```


\noindent 6. Split hospitals into quartiles based on bed size and provide a table of the average price among treated/control groups for each quartile.

```{r}
#| echo: false
#| label: tbl-plans
#| tbl-cap: Multiple reports

knitr::kable(
  quartile_summary,
  col.names = c("Bed Quartile", "No Penalty", "Penalty"),
  caption = "Average Prices by Bed Quartile and Penalty Status",
  format = "markdown"
)
```


\noindent 7. Find the average treatment effect using each of the following estimators, and present your results in a single table:
a. Nearest neighbor matching (1-to-1) with inverse variance distance based on quartiles of bed size
b. Nearest neighbor matching (1-to-1) with Mahalanobis distance based on quartiles of bed size
c. Inverse propensity weighting, where the propensity scores are based on quartiles of bed size
d. Simple linear regression, adjusting for quartiles of bed size using dummy variables and appropriate interactions as discussed in class

```{r}
#| echo: false
#| label: tbl-plans
#| tbl-cap: Multiple reports

# Extract ATE estimates
ate_near_match <- near.match$est
ate_mal_match <- mal.match$est
ate_ipw_diff <- ipw.diff
ate_regression <- ate

# Create a data frame summarizing the results
ate_estimates <- data.frame(
  Method = c(
    "Nearest Matching (Inverse Variance)",
    "Nearest Matching (Mahalanobis Distance)",
    "Inverse Propensity Weighting (IPW)",
    "Linear Regression"
  ),
  ATE_Estimate = c(ate_near_match, ate_mal_match, ate_ipw_diff, ate_regression)
)

# Print the table using knitr::kable()
knitr::kable(ate_estimates, 
             caption = "ATE Estimates from Different Methods", 
             col.names = c("Method", "ATE Estimate"))
```